### 2010-08-09 Thesis/Chapter10.tex
- Update 9 August -
Added all chapters files + transition chapter2-chapter3 + summary chapter3 + removed outline



git-svn-id: svn+ssh://scm.gforge.inria.fr/svnroot/sofa-shells@146 7b2761c9-2758-49ef-8c29-9c9f1e1aad1b

\part{Introduction}

	\chapter{Medical simulation}
		\section{General context}
			\subsection{Simulators for everything: we started by simulating what is simple}
			\subsection{As complexity increases, applications to medical training, patient-specific planning and per-operative guidance}
		\section{Challenges}					
			\subsection{Photorealistic rendering}
			\subsection{Accurate haptic feedback}
			\subsection{Physically realist organ modelling in real-time}

	\input{chapter2}

	\input{chapter3}

\part{Solid organs modelling}

	\chapter{State of art: FEM}
		\section{Expose the problem: how to model solid organs?}
		\section{Techniques using geometrical kind of approaches}
			\subsection{Shape matching}
		\section{Techniques relying on physics}
			\subsection{Mass-spring}
		\section{Techniques based on continuum mechanics}	
			\subsection{FEM with mesh (highlight evolution: linear, topological changes, co-rotational, non-linear, GPU)}
			\subsection{Meshless (supposedly good for handling topological changes but other issues}

	\chapter{Introduction of TLED}
		\section{Description of the algorithm (neo-Hookean, explicit integration, independent elements, parallelisable on GPU)}
		\section{Visco-elasticity and anisotropy added \OC{MICCAI 2008; MedIA 2009}}

	\chapter{GPU implementation of TLED}
		\section{What is GPGPU}
		\section{Re-formulation of the algorithm for its Cg implementation}
		\section{CUDA implementation/optimisations \OC{ISBMS 2008a}}
		\section{Implementation into SOFA}
			\subsection{Presentation of SOFA project and architecture}
			\subsection{TLED in SOFA and release in open-source}

\part{Hollow organs modelling}

	\chapter{State of art: hollow structures}
		\section{Expose the problem: models presented for solid organs are insufficient. How to model hollow organs?}
			\subsection{Colonoscopy simulator project: needs for colon}
			\subsection{Cataract surgery, stenosis: other needs for implants/blood vessels}
		\section{Non-physic approaches (computer graphics stuff)}
		\section{Physically accurate approches (plates/shells)}

	\chapter{Mass-spring model for colon implemented on GPU \OC{ISBMS 2008b}}
		\section{Colon is big, we need a fast model}
		\section{Limitations (difficult to link it to actual physical properties)}

	\chapter{A co-rotational triangular shell model \OC{ISBMS 2010}}
		\section{Model description}
		\section{Validation}
		\section{Application to implant deployment simulation in cataract surgery}

	\chapter{'Shell meshing' technique \OC{MICCAI 2010}}
		\section{State of art: reconstruction/simplification}
		\section{Our method}

\part{Conclusion}
	\chapter{Summary}
	\chapter{Discussion and perspectives (Interaction solid/hollow organs)}

>>>>

\chapter{'Shell meshing' technique \OC{MICCAI 2010}}
\label{chap10}
\begin{shortAbstract}
A short abstract for the upcoming chapter
\end{shortAbstract}

\section{State of art: reconstruction/simplification}

\section{Our method}

### 2010-08-09 Thesis/Chapter2.tex
- Update 30 July - Changed outline according to what we discussed with Stephane + increased level of details + bib updated + creation of chapter 3
git-svn-id: svn+ssh://scm.gforge.inria.fr/svnroot/sofa-shells@138 7b2761c9-2758-49ef-8c29-9c9f1e1aad1b
- Update 2 August - Started chapter3 + minor changes on chapter2 + bib updated
git-svn-id: svn+ssh://scm.gforge.inria.fr/svnroot/sofa-shells@139 7b2761c9-2758-49ef-8c29-9c9f1e1aad1b
- Update 9 August -
Added all chapters files + transition chapter2-chapter3 + summary chapter3 + removed outline



git-svn-id: svn+ssh://scm.gforge.inria.fr/svnroot/so

\chapter{Background in continuum mechanics for soft-tissue modelling}
\label{chap:softtissue}
\begin{shortAbstract}
As seen in the previous chapter, realistic modelling of organ deformation is a challenging research field that opens the door to new clinical applications including: medical training and rehearsal systems, patient-specific planning of surgical procedures and per-operative guidance based on simulation. In all these cases the clinician needs fast updates of the deformation model to obtain a real-time display of the computed deformations. If for medical training devices the haptic feedback from touching organs merely needs to feel real, the accuracy of the information provided to the clinician in the cases of planning or per-operative guidance is crucial. Therefore, a substantial comprehension of the mechanics involved and a knowledge of the physical properties of anatomical structures are both mandatory in our quest to realistically model the deformation of organs. This chapter will start by introducing the main concepts of continuum mechanics that are fundamental to study the mechanical response of organs. It will then present mathematical models able to describe the different mechanical aspects of materials and briefly discuss the mechanical characterisation of tissues.
\end{shortAbstract}

@@

	\subsection{Isotropic materials} \label{chap2:isotropic}
A material is said \emph{isotropic} if its properties are independent of directions. Therefore, we have the following relations:
\begin{equation}
E_1 = E_2 = E_3 = E, \quad G_{12} = G_{13} = G_{23} = G, \quad \nu_{12} = \nu_{23} = \nu_{13} = \nu.
\end{equation}
The stress-strain relations may be written:
\begin{equation}
\sigma_{ij} = \frac{E}{1+\nu} \varepsilon_{ij} + \frac{\nu E}{(1+\nu)(1-2\nu)} \varepsilon_{kk}\delta_{ij}.
\end{equation}	
where the summation of repeated indices is implied (Einstein's notation). The inverse relations are:
\begin{equation}
\varepsilon_{ij} = \frac{1+\nu}{E} \sigma_{ij} - \frac{\nu}{E} \sigma_{kk} \delta_{ij},
\end{equation}
recalling that $G = E/[2(1+\nu)]$.

@@

@@

\TODO{Transition with next chapter (complexity of models => FEM to solve them)}

>>>>

\chapter{Background in continuum mechanics for soft-tissue modelling}
\label{chap2}
\begin{shortAbstract}
As seen in the previous chapter, realistic modelling of organ deformation is a challenging research field that opens the door to new clinical applications including: medical training and rehearsal systems, patient-specific planning of surgical procedures and per-operative guidance based on simulation. In all these cases the clinician needs fast updates of the deformation model to obtain a real-time display of the computed deformations. If for medical training devices the haptic feedback from touching organs merely needs to feel real, the accuracy of the information provided to the clinician in the cases of planning or per-operative guidance is crucial. Therefore, a substantial comprehension of the mechanics involved and a knowledge of the physical properties of anatomical structures are both mandatory in our quest to realistically model the deformation of organs. This chapter will start by introducing the main concepts of continuum mechanics that are fundamental to study the mechanical response of organs. It will then present mathematical models able to describe the different mechanical aspects of materials and briefly discuss the mechanical characterisation of tissues.
\end{shortAbstract}

@@

	\subsection{Isotropic materials} \label{chap2:isotropic}
A material is said \emph{isotropic} if its properties are independent of directions. Therefore, we have the following relations:
\begin{equation}
E_1 = E_2 = E_3 = E, \quad G_{12} = G_{13} = G_{23} = G, \quad \nu_{12} = \nu_{23} = \nu_{13} = \nu.
\end{equation}
The stress-strain relations may be written:
\begin{equation}
\sigma_{ij} = \frac{E}{1+\nu} \varepsilon_{ij} + \frac{\nu E}{(1+\nu)(1-2\nu)} \varepsilon_{kk}\delta_{ij}.
\end{equation}	
where the summation of repeated indices is implied (Einstein's notation) and $\delta_{ij}$ is the Kronecker delta\footnote{$ \delta_{ij}$ is equal to $1$ if $i = j$ and to $0$ if $i \neq j$}. The inverse relations are:
\begin{equation}
\varepsilon_{ij} = \frac{1+\nu}{E} \sigma_{ij} - \frac{\nu}{E} \sigma_{kk} \delta_{ij},
\end{equation}
recalling that $G = E/[2(1+\nu)]$.

@@

In other words, the resistance of a viscoelastic material to a given force depends on the velocity to which this force is applied.

@@

These various mechanical aspects of materials make realistic modelling of anatomical structures fairly challenging. In medical simulation, where modelling must occur in real-time or close to real-time, very often simplifications have to be made. As we will see in chapter \ref{chap4} and \ref{chap7}, numerous models are available in the literature to describe physics of objects, from fairly simple and naive approaches to more complex and thorough representations. However, the most common technique to solve equations of continuum mechanics is the Finite Element Method. Moreover, since the two main contributions of this thesis deal with finite element methods formulations, the general principles of this method will be detailled in the next section.

### 2010-08-09 Thesis/Chapter3.tex
- Update 9 August -
Added all chapters files + transition chapter2-chapter3 + summary chapter3 + removed outline



git-svn-id: svn+ssh://scm.gforge.inria.fr/svnroot/sofa-shells@146 7b2761c9-2758-49ef-8c29-9c9f1e1aad1b

\chapter{Practical approach of the Finite Element Method}
\label{chap3}
\begin{shortAbstract}
Blabla
\end{shortAbstract}

@@

	\subsection{Time dependence}
In all the equations of the finite element method that we derived so far, we have not taken time into account. For some problem, this is not an issue, the motion is sufficiently slow to assume that the dynamic effects (such as damping effect) are neglectable. However, this assumption is not always valid, particulary in our case where we want to model organ deformation. The deformation obviously depends on time. Finite element models of time-dependent problems can be developed in two alternatives ways \citep{Reddy93}:
\begin{description}
\item[(a)] coupled formulation in which the time $ t $ is treated as an additional coordinate along with the spatial coordinate $ x $
\item[(b)] decoupled formulation where time and spatial variations are assumede to be separable.
\end{description}
Thus, the approximation of the solution $ u $ takes one of these two forms:
\begin{align}
u(x, t) \approx u^e(x, t) &= \sum_{i=1}^n \hat{u}^e_i \hat{\psi}^e_i(x, t)  \quad \text{ (coupled formulation)} \label{chap3:approxTime1}\\
u(x, t) \approx u^e(x, t) &= \sum_{i=1}^n u^e_i(t) \psi^e_i(x) \quad \text{(decoupled formulation)} \label{chap3:approxTime2}
\end{align}
where $ \hat{\psi}^e_i(x, t) $ are time-space interpolation functions, $ \hat{u}^e_i $ are the nodal values that are independent of $ x $ and $ t $, $ \psi^e_i(x) $ are the usual interpolation functions in spatial coordinate $ x $ only and the nodal values $ u^e_i(t) $ are functions of time $ t $ only. Space-time coupled finite element formulations are not common and they have not been adequately studied. Hence, we consider the decoupled formulation only. Of course, the assumption that the time and spatial variations are separable is not valid in general. However, with sufficiently small time steps, it is possible to obtain accurate solutions nevertheless.

@@

The construction of the weak form using either energy principles or weighted residual methods will not be detailled here. The reader may easily find it in dozens of textbooks. By substituting \eqref{chap3:approxTime} in the weak form, we obtain the finite element equations of equilibrium in matrix form:
\begin{equation}
\label{chap3:eqDynamic}
\mathbf{M} \mathbf{\ddot U} + \mathbf{D} \mathbf{ \dot U} + \mathbf{K}(\mathbf{U}) \cdot \mathbf{U} = \mathbf{R},
\end{equation}
where $ \mathbf{M} $ is a constant mass matrix, $\mathbf{D}$ is a constant damping matrix, $ \mathbf{K}(\mathbf{U}) $ is the stiffness matrix, which is a function of global nodal displacements $\mathbf{U}$, and $\mathbf{R}$ are externally applied loads. This equation is applicable to individual elements as well as the entire element assemblage.

>>>>

\chapter{Practical approach of the Finite Element Method}
\label{chap3}
\begin{shortAbstract}
The previous chapter explained in details how physics of anatomical structures could be described by continuun mechanics. When an organ is submitted to a force, the first effect of this application is a change in the organ's shape. This deformation can be measured using different strain tensors. By modelling the mechanical aspects of the material with an appropriate constitutive law, we can then deduce the stress tensor (or internal forces) at each point within the organ. The next step is to apply Newton's second law (law of conservation of momentum), which yields a partial differential equation at every point of the volume. The Finite Element Method is a numerical procedure that allows to turn an infinite number of partial differential equations into a finite number of algebraic equations. The method follows different steps. We will first begin by describing the subdivision of the domain into smaller sub-domains into which local simpler equations are derived. We will then discuss how all the local equations are summed over the whole domain which eventually leads to the global solution. 
\end{shortAbstract}

@@

	\subsection{Time dependence}
In all the equations of the finite element method that we derived so far, we have not taken time into account. For some problem, this is not an issue, the motion is sufficiently slow to assume that the dynamic effects (such as damping effect) are neglectable. However, this assumption is not always valid, particulary in our case where we want to model organ deformation. The deformation obviously depends on time. Finite element models of time-dependent problems can be developed in two alternatives ways \citep{Reddy93}:
\begin{description}
\item[(a)] coupled formulation in which the time $ t $ is treated as an additional coordinate along with the spatial coordinate $ x $
\item[(b)] decoupled formulation where time and spatial variations are assumed to be separable.
\end{description}
Thus, the approximation of the solution $ u $ takes one of these two forms:
\begin{align}
u(x, t) \approx u^e(x, t) &= \sum_{i=1}^n \hat{u}^e_i \hat{\psi}^e_i(x, t)  \quad \text{ (coupled formulation)} \label{chap3:approxTime1}\\
u(x, t) \approx u^e(x, t) &= \sum_{i=1}^n u^e_i(t) \psi^e_i(x) \quad \text{(decoupled formulation)} \label{chap3:approxTime2}
\end{align}
where $ \hat{\psi}^e_i(x, t) $ are time-space interpolation functions, $ \hat{u}^e_i $ are the nodal values that are independent of $ x $ and $ t $, $ \psi^e_i(x) $ are the usual interpolation functions in spatial coordinate $ x $ only and the nodal values $ u^e_i(t) $ are functions of time $ t $ only. Space-time coupled finite element formulations are not common and they have not been adequately studied. Hence, we consider the decoupled formulation only. Of course, the assumption that the time and spatial variations are separable is not valid in general. However, with sufficiently small time steps, it is possible to obtain accurate solutions nevertheless.

@@

The construction of the weak form using either energy principles or weighted residual methods will not be detailled here. The reader may easily find it in dozens of textbooks. By substituting \eqref{chap3:approxTime} in the weak form, we obtain the finite element equations of equilibrium in matrix form:
\begin{equation}
\label{chap3:eqDynamic}
\mathbf{M} \mathbf{\ddot U} + \mathbf{D} \mathbf{ \dot U} + \mathbf{K}(\mathbf{U}) \cdot \mathbf{U} = \mathbf{R},
\end{equation}
where $ \mathbf{M} $ is a constant mass matrix, $\mathbf{D}$ is a constant damping matrix, $ \mathbf{K}(\mathbf{U}) $ is the stiffness matrix, which is a function of global nodal displacements $\mathbf{U}$, and $\mathbf{R}$ are externally applied loads. This equation is applicable to individual elements as well as the entire element assemblage (this is why the subscripts $ ^e $ were dropped).

### 2010-08-09 Thesis/Chapter4.tex
- Update 9 August -
Added all chapters files + transition chapter2-chapter3 + summary chapter3 + removed outline



git-svn-id: svn+ssh://scm.gforge.inria.fr/svnroot/sofa-shells@146 7b2761c9-2758-49ef-8c29-9c9f1e1aad1b

\part{Introduction}

	\chapter{Medical simulation}
		\section{General context}
			\subsection{Simulators for everything: we started by simulating what is simple}
			\subsection{As complexity increases, applications to medical training, patient-specific planning and per-operative guidance}
		\section{Challenges}					
			\subsection{Photorealistic rendering}
			\subsection{Accurate haptic feedback}
			\subsection{Physically realist organ modelling in real-time}

	\input{chapter2}

	\input{chapter3}

\part{Solid organs modelling}

	\chapter{State of art: FEM}
		\section{Expose the problem: how to model solid organs?}
		\section{Techniques using geometrical kind of approaches}
			\subsection{Shape matching}
		\section{Techniques relying on physics}
			\subsection{Mass-spring}
		\section{Techniques based on continuum mechanics}	
			\subsection{FEM with mesh (highlight evolution: linear, topological changes, co-rotational, non-linear, GPU)}
			\subsection{Meshless (supposedly good for handling topological changes but other issues}

	\chapter{Introduction of TLED}
		\section{Description of the algorithm (neo-Hookean, explicit integration, independent elements, parallelisable on GPU)}
		\section{Visco-elasticity and anisotropy added \OC{MICCAI 2008; MedIA 2009}}

	\chapter{GPU implementation of TLED}
		\section{What is GPGPU}
		\section{Re-formulation of the algorithm for its Cg implementation}
		\section{CUDA implementation/optimisations \OC{ISBMS 2008a}}
		\section{Implementation into SOFA}
			\subsection{Presentation of SOFA project and architecture}
			\subsection{TLED in SOFA and release in open-source}

\part{Hollow organs modelling}

	\chapter{State of art: hollow structures}
		\section{Expose the problem: models presented for solid organs are insufficient. How to model hollow organs?}
			\subsection{Colonoscopy simulator project: needs for colon}
			\subsection{Cataract surgery, stenosis: other needs for implants/blood vessels}
		\section{Non-physic approaches (computer graphics stuff)}
		\section{Physically accurate approches (plates/shells)}

	\chapter{Mass-spring model for colon implemented on GPU \OC{ISBMS 2008b}}
		\section{Colon is big, we need a fast model}
		\section{Limitations (difficult to link it to actual physical properties)}

	\chapter{A co-rotational triangular shell model \OC{ISBMS 2010}}
		\section{Model description}
		\section{Validation}
		\section{Application to implant deployment simulation in cataract surgery}

	\chapter{'Shell meshing' technique \OC{MICCAI 2010}}
		\section{State of art: reconstruction/simplification}
		\section{Our method}

\part{Conclusion}
	\chapter{Summary}
	\chapter{Discussion and perspectives (Interaction solid/hollow organs)}

>>>>

\chapter{State of art: FEM}
\label{chap4}
\begin{shortAbstract}
A short abstract for the upcoming chapter
\end{shortAbstract}

\section{Expose the problem: how to model solid organs?}

\section{Techniques using geometrical kind of approaches}
	\subsection{Shape matching}
	\section{Techniques relying on physics}
	\subsection{Mass-spring}

\section{Techniques based on continuum mechanics}	
	\subsection{FEM with mesh (highlight evolution: linear, topological changes, co-rotational, non-linear, GPU)}
	\subsection{Meshless (supposedly good for handling topological changes but other issues}

### 2010-08-09 Thesis/Chapter5.tex
- Update 9 August -
Added all chapters files + transition chapter2-chapter3 + summary chapter3 + removed outline



git-svn-id: svn+ssh://scm.gforge.inria.fr/svnroot/sofa-shells@146 7b2761c9-2758-49ef-8c29-9c9f1e1aad1b

\part{Introduction}

	\chapter{Medical simulation}
		\section{General context}
			\subsection{Simulators for everything: we started by simulating what is simple}
			\subsection{As complexity increases, applications to medical training, patient-specific planning and per-operative guidance}
		\section{Challenges}					
			\subsection{Photorealistic rendering}
			\subsection{Accurate haptic feedback}
			\subsection{Physically realist organ modelling in real-time}

	\input{chapter2}

	\input{chapter3}

\part{Solid organs modelling}

	\chapter{State of art: FEM}
		\section{Expose the problem: how to model solid organs?}
		\section{Techniques using geometrical kind of approaches}
			\subsection{Shape matching}
		\section{Techniques relying on physics}
			\subsection{Mass-spring}
		\section{Techniques based on continuum mechanics}	
			\subsection{FEM with mesh (highlight evolution: linear, topological changes, co-rotational, non-linear, GPU)}
			\subsection{Meshless (supposedly good for handling topological changes but other issues}

	\chapter{Introduction of TLED}
		\section{Description of the algorithm (neo-Hookean, explicit integration, independent elements, parallelisable on GPU)}
		\section{Visco-elasticity and anisotropy added \OC{MICCAI 2008; MedIA 2009}}

	\chapter{GPU implementation of TLED}
		\section{What is GPGPU}
		\section{Re-formulation of the algorithm for its Cg implementation}
		\section{CUDA implementation/optimisations \OC{ISBMS 2008a}}
		\section{Implementation into SOFA}
			\subsection{Presentation of SOFA project and architecture}
			\subsection{TLED in SOFA and release in open-source}

\part{Hollow organs modelling}

	\chapter{State of art: hollow structures}
		\section{Expose the problem: models presented for solid organs are insufficient. How to model hollow organs?}
			\subsection{Colonoscopy simulator project: needs for colon}
			\subsection{Cataract surgery, stenosis: other needs for implants/blood vessels}
		\section{Non-physic approaches (computer graphics stuff)}
		\section{Physically accurate approches (plates/shells)}

	\chapter{Mass-spring model for colon implemented on GPU \OC{ISBMS 2008b}}
		\section{Colon is big, we need a fast model}
		\section{Limitations (difficult to link it to actual physical properties)}

	\chapter{A co-rotational triangular shell model \OC{ISBMS 2010}}
		\section{Model description}
		\section{Validation}
		\section{Application to implant deployment simulation in cataract surgery}

	\chapter{'Shell meshing' technique \OC{MICCAI 2010}}
		\section{State of art: reconstruction/simplification}
		\section{Our method}

\part{Conclusion}
	\chapter{Summary}
	\chapter{Discussion and perspectives (Interaction solid/hollow organs)}

>>>>

\chapter{Introduction of TLED}
\label{chap5}
\begin{shortAbstract}
A short abstract for the upcoming chapter
\end{shortAbstract}

\section{Description of the algorithm (neo-Hookean, explicit integration, independent elements, parallelisable on GPU)}

\section{Visco-elasticity and anisotropy added \OC{MICCAI 2008; MedIA 2009}}

### 2010-08-09 Thesis/Chapter6.tex
- Update 9 August -
Added all chapters files + transition chapter2-chapter3 + summary chapter3 + removed outline



git-svn-id: svn+ssh://scm.gforge.inria.fr/svnroot/sofa-shells@146 7b2761c9-2758-49ef-8c29-9c9f1e1aad1b

\part{Introduction}

	\chapter{Medical simulation}
		\section{General context}
			\subsection{Simulators for everything: we started by simulating what is simple}
			\subsection{As complexity increases, applications to medical training, patient-specific planning and per-operative guidance}
		\section{Challenges}					
			\subsection{Photorealistic rendering}
			\subsection{Accurate haptic feedback}
			\subsection{Physically realist organ modelling in real-time}

	\input{chapter2}

	\input{chapter3}

\part{Solid organs modelling}

	\chapter{State of art: FEM}
		\section{Expose the problem: how to model solid organs?}
		\section{Techniques using geometrical kind of approaches}
			\subsection{Shape matching}
		\section{Techniques relying on physics}
			\subsection{Mass-spring}
		\section{Techniques based on continuum mechanics}	
			\subsection{FEM with mesh (highlight evolution: linear, topological changes, co-rotational, non-linear, GPU)}
			\subsection{Meshless (supposedly good for handling topological changes but other issues}

	\chapter{Introduction of TLED}
		\section{Description of the algorithm (neo-Hookean, explicit integration, independent elements, parallelisable on GPU)}
		\section{Visco-elasticity and anisotropy added \OC{MICCAI 2008; MedIA 2009}}

	\chapter{GPU implementation of TLED}
		\section{What is GPGPU}
		\section{Re-formulation of the algorithm for its Cg implementation}
		\section{CUDA implementation/optimisations \OC{ISBMS 2008a}}
		\section{Implementation into SOFA}
			\subsection{Presentation of SOFA project and architecture}
			\subsection{TLED in SOFA and release in open-source}

\part{Hollow organs modelling}

	\chapter{State of art: hollow structures}
		\section{Expose the problem: models presented for solid organs are insufficient. How to model hollow organs?}
			\subsection{Colonoscopy simulator project: needs for colon}
			\subsection{Cataract surgery, stenosis: other needs for implants/blood vessels}
		\section{Non-physic approaches (computer graphics stuff)}
		\section{Physically accurate approches (plates/shells)}

	\chapter{Mass-spring model for colon implemented on GPU \OC{ISBMS 2008b}}
		\section{Colon is big, we need a fast model}
		\section{Limitations (difficult to link it to actual physical properties)}

	\chapter{A co-rotational triangular shell model \OC{ISBMS 2010}}
		\section{Model description}
		\section{Validation}
		\section{Application to implant deployment simulation in cataract surgery}

	\chapter{'Shell meshing' technique \OC{MICCAI 2010}}
		\section{State of art: reconstruction/simplification}
		\section{Our method}

\part{Conclusion}
	\chapter{Summary}
	\chapter{Discussion and perspectives (Interaction solid/hollow organs)}

>>>>

\chapter{GPU implementation of TLED}
\label{chap6}
\begin{shortAbstract}
A short abstract for the upcoming chapter
\end{shortAbstract}

\section{What is GPGPU}

\section{Re-formulation of the algorithm for its Cg implementation}

\section{CUDA implementation/optimisations \OC{ISBMS 2008a}}

\section{Implementation into SOFA}
	\subsection{Presentation of SOFA project and architecture}
	\subsection{TLED in SOFA and release in open-source}

### 2010-08-09 Thesis/Chapter7.tex
- Update 9 August -
Added all chapters files + transition chapter2-chapter3 + summary chapter3 + removed outline



git-svn-id: svn+ssh://scm.gforge.inria.fr/svnroot/sofa-shells@146 7b2761c9-2758-49ef-8c29-9c9f1e1aad1b

\part{Introduction}

	\chapter{Medical simulation}
		\section{General context}
			\subsection{Simulators for everything: we started by simulating what is simple}
			\subsection{As complexity increases, applications to medical training, patient-specific planning and per-operative guidance}
		\section{Challenges}					
			\subsection{Photorealistic rendering}
			\subsection{Accurate haptic feedback}
			\subsection{Physically realist organ modelling in real-time}

	\input{chapter2}

	\input{chapter3}

\part{Solid organs modelling}

	\chapter{State of art: FEM}
		\section{Expose the problem: how to model solid organs?}
		\section{Techniques using geometrical kind of approaches}
			\subsection{Shape matching}
		\section{Techniques relying on physics}
			\subsection{Mass-spring}
		\section{Techniques based on continuum mechanics}	
			\subsection{FEM with mesh (highlight evolution: linear, topological changes, co-rotational, non-linear, GPU)}
			\subsection{Meshless (supposedly good for handling topological changes but other issues}

	\chapter{Introduction of TLED}
		\section{Description of the algorithm (neo-Hookean, explicit integration, independent elements, parallelisable on GPU)}
		\section{Visco-elasticity and anisotropy added \OC{MICCAI 2008; MedIA 2009}}

	\chapter{GPU implementation of TLED}
		\section{What is GPGPU}
		\section{Re-formulation of the algorithm for its Cg implementation}
		\section{CUDA implementation/optimisations \OC{ISBMS 2008a}}
		\section{Implementation into SOFA}
			\subsection{Presentation of SOFA project and architecture}
			\subsection{TLED in SOFA and release in open-source}

\part{Hollow organs modelling}

	\chapter{State of art: hollow structures}
		\section{Expose the problem: models presented for solid organs are insufficient. How to model hollow organs?}
			\subsection{Colonoscopy simulator project: needs for colon}
			\subsection{Cataract surgery, stenosis: other needs for implants/blood vessels}
		\section{Non-physic approaches (computer graphics stuff)}
		\section{Physically accurate approches (plates/shells)}

	\chapter{Mass-spring model for colon implemented on GPU \OC{ISBMS 2008b}}
		\section{Colon is big, we need a fast model}
		\section{Limitations (difficult to link it to actual physical properties)}

	\chapter{A co-rotational triangular shell model \OC{ISBMS 2010}}
		\section{Model description}
		\section{Validation}
		\section{Application to implant deployment simulation in cataract surgery}

	\chapter{'Shell meshing' technique \OC{MICCAI 2010}}
		\section{State of art: reconstruction/simplification}
		\section{Our method}

\part{Conclusion}
	\chapter{Summary}
	\chapter{Discussion and perspectives (Interaction solid/hollow organs)}

>>>>

\chapter{State of art: hollow structures}
\label{chap7}
\begin{shortAbstract}
A short abstract for the upcoming chapter
\end{shortAbstract}

\section{Expose the problem: models presented for solid organs are insufficient. How to model hollow organs?}
	\subsection{Colonoscopy simulator project: needs for colon}
	\subsection{Cataract surgery, stenosis: other needs for implants/blood vessels}

\section{Non-physic approaches (computer graphics stuff)}

\section{Physically accurate approches (plates/shells)}

### 2010-08-09 Thesis/Chapter8.tex
- Update 9 August -
Added all chapters files + transition chapter2-chapter3 + summary chapter3 + removed outline



git-svn-id: svn+ssh://scm.gforge.inria.fr/svnroot/sofa-shells@146 7b2761c9-2758-49ef-8c29-9c9f1e1aad1b

\part{Introduction}

	\chapter{Medical simulation}
		\section{General context}
			\subsection{Simulators for everything: we started by simulating what is simple}
			\subsection{As complexity increases, applications to medical training, patient-specific planning and per-operative guidance}
		\section{Challenges}					
			\subsection{Photorealistic rendering}
			\subsection{Accurate haptic feedback}
			\subsection{Physically realist organ modelling in real-time}

	\input{chapter2}

	\input{chapter3}

\part{Solid organs modelling}

	\chapter{State of art: FEM}
		\section{Expose the problem: how to model solid organs?}
		\section{Techniques using geometrical kind of approaches}
			\subsection{Shape matching}
		\section{Techniques relying on physics}
			\subsection{Mass-spring}
		\section{Techniques based on continuum mechanics}	
			\subsection{FEM with mesh (highlight evolution: linear, topological changes, co-rotational, non-linear, GPU)}
			\subsection{Meshless (supposedly good for handling topological changes but other issues}

	\chapter{Introduction of TLED}
		\section{Description of the algorithm (neo-Hookean, explicit integration, independent elements, parallelisable on GPU)}
		\section{Visco-elasticity and anisotropy added \OC{MICCAI 2008; MedIA 2009}}

	\chapter{GPU implementation of TLED}
		\section{What is GPGPU}
		\section{Re-formulation of the algorithm for its Cg implementation}
		\section{CUDA implementation/optimisations \OC{ISBMS 2008a}}
		\section{Implementation into SOFA}
			\subsection{Presentation of SOFA project and architecture}
			\subsection{TLED in SOFA and release in open-source}

\part{Hollow organs modelling}

	\chapter{State of art: hollow structures}
		\section{Expose the problem: models presented for solid organs are insufficient. How to model hollow organs?}
			\subsection{Colonoscopy simulator project: needs for colon}
			\subsection{Cataract surgery, stenosis: other needs for implants/blood vessels}
		\section{Non-physic approaches (computer graphics stuff)}
		\section{Physically accurate approches (plates/shells)}

	\chapter{Mass-spring model for colon implemented on GPU \OC{ISBMS 2008b}}
		\section{Colon is big, we need a fast model}
		\section{Limitations (difficult to link it to actual physical properties)}

	\chapter{A co-rotational triangular shell model \OC{ISBMS 2010}}
		\section{Model description}
		\section{Validation}
		\section{Application to implant deployment simulation in cataract surgery}

	\chapter{'Shell meshing' technique \OC{MICCAI 2010}}
		\section{State of art: reconstruction/simplification}
		\section{Our method}

\part{Conclusion}
	\chapter{Summary}
	\chapter{Discussion and perspectives (Interaction solid/hollow organs)}

>>>>

\chapter{Mass-spring model for colon implemented on GPU \OC{ISBMS 2008b}}
\label{chap8}
\begin{shortAbstract}
A short abstract for the upcoming chapter
\end{shortAbstract}

\section{Colon is big, we need a fast model}

\section{Limitations (difficult to link it to actual physical properties)}

### 2010-08-09 Thesis/Chapter9.tex
- Update 9 August -
Added all chapters files + transition chapter2-chapter3 + summary chapter3 + removed outline



git-svn-id: svn+ssh://scm.gforge.inria.fr/svnroot/sofa-shells@146 7b2761c9-2758-49ef-8c29-9c9f1e1aad1b

\part{Introduction}

	\chapter{Medical simulation}
		\section{General context}
			\subsection{Simulators for everything: we started by simulating what is simple}
			\subsection{As complexity increases, applications to medical training, patient-specific planning and per-operative guidance}
		\section{Challenges}					
			\subsection{Photorealistic rendering}
			\subsection{Accurate haptic feedback}
			\subsection{Physically realist organ modelling in real-time}

	\input{chapter2}

	\input{chapter3}

\part{Solid organs modelling}

	\chapter{State of art: FEM}
		\section{Expose the problem: how to model solid organs?}
		\section{Techniques using geometrical kind of approaches}
			\subsection{Shape matching}
		\section{Techniques relying on physics}
			\subsection{Mass-spring}
		\section{Techniques based on continuum mechanics}	
			\subsection{FEM with mesh (highlight evolution: linear, topological changes, co-rotational, non-linear, GPU)}
			\subsection{Meshless (supposedly good for handling topological changes but other issues}

	\chapter{Introduction of TLED}
		\section{Description of the algorithm (neo-Hookean, explicit integration, independent elements, parallelisable on GPU)}
		\section{Visco-elasticity and anisotropy added \OC{MICCAI 2008; MedIA 2009}}

	\chapter{GPU implementation of TLED}
		\section{What is GPGPU}
		\section{Re-formulation of the algorithm for its Cg implementation}
		\section{CUDA implementation/optimisations \OC{ISBMS 2008a}}
		\section{Implementation into SOFA}
			\subsection{Presentation of SOFA project and architecture}
			\subsection{TLED in SOFA and release in open-source}

\part{Hollow organs modelling}

	\chapter{State of art: hollow structures}
		\section{Expose the problem: models presented for solid organs are insufficient. How to model hollow organs?}
			\subsection{Colonoscopy simulator project: needs for colon}
			\subsection{Cataract surgery, stenosis: other needs for implants/blood vessels}
		\section{Non-physic approaches (computer graphics stuff)}
		\section{Physically accurate approches (plates/shells)}

	\chapter{Mass-spring model for colon implemented on GPU \OC{ISBMS 2008b}}
		\section{Colon is big, we need a fast model}
		\section{Limitations (difficult to link it to actual physical properties)}

	\chapter{A co-rotational triangular shell model \OC{ISBMS 2010}}
		\section{Model description}
		\section{Validation}
		\section{Application to implant deployment simulation in cataract surgery}

	\chapter{'Shell meshing' technique \OC{MICCAI 2010}}
		\section{State of art: reconstruction/simplification}
		\section{Our method}

\part{Conclusion}
	\chapter{Summary}
	\chapter{Discussion and perspectives (Interaction solid/hollow organs)}

>>>>

\chapter{A co-rotational triangular shell model \OC{ISBMS 2010}}
\label{chap9}
\begin{shortAbstract}
A short abstract for the upcoming chapter
\end{shortAbstract}

\section{Model description}

\section{Validation}

\section{Application to implant deployment simulation in cataract surgery}
